\chapter{Related work}
\label{ch:related-work}
This chapter reviews the work related to improving Korat and static analysis. Section \ref{sec:attempts-to-improve-korat} discusses other projects that have improved different aspects of Korat. Section \ref{sec:related-static-analyses} discusses a few static program analysis tools that have tried to verify properties of data structures.

\section{Attemps to improve Korat}
\label{sec:attempts-to-improve-korat}
Over the last decade, several projects have tried to improve various
aspects of Korat. Parallel test generation and execution with Korat
\cite{misailovic2007parallel} talks about techniques to generate test inputs with a
parallel search in Korat and techniques to execute test inputs in
parallel, both off-line (when the inputs are saved on disk) and
on-line (when execution immediately follows generation).PKorat
\cite{siddiqui2009pkorat} improves on the previous work and explores the same state
space as Korat but considers several candidates in each
iteration. These candidates are distributed among parallel workers
resulting in an efficient parallel version of Korat.



\section{Static Analysis}
\label{sec:related-static-analyses}
Several projects have developed static analyses for verifying program
properties. We will discuss some of the projects that check properties
of complex data structures. The Extended Static Checker (ESC)
\cite{flanagan2002extended} uses a theorem prover to verify partial correctness of Java
classes annotated with JML specifications. ESC can find faults such as
null pointer dereferences, array bounds violations and division by
zero. The Three-Valued-Logic Analyzer (TVLA) \cite{sagiv2002parametric} was the first
system to implement a shape analysis, which verifies that programs
with destructive updates preserve structural invariants of linked data
structures. There has been work \cite{yorsh2004symbolically}  that extends TVLA to
handle more properties of the content of the data structures and uses
symbolic computation to automatically derive transfer functions, thus
reducing the amount of work that the user needs to do.

\para Kuncak et al. proposed Role Analysis \cite{kuncak2002role}, a compositional
interprocedual analysis that verifies similar properties like
TVLA. The same authors also developed Hob \cite{lam2005generalized}, an analysis
framework that combines different static analysis and theorem proving
techniques \cite{zee2004combining} to show complex data structure consistency
properties. The Pointer Assertion Logic Engine (PALE) \cite{moller2001pointer} can
verify a large set of data structures that have a spanning tree
backbone, with possibly additional pointers that do not add extra
information. Jalloy \cite{jackson2000finding,vaziri2003checking} analyzes methods that manipulate linked data
structures by first building an alloy model of Java code and then
checking it in a bounded-exhaustive manner with the Alloy Analyzer
\cite{jackson2000alcoa}.


